---
title: "R Notebook"
output: html_notebook
---

Uses data from bycaught olive ridlesy in the central North Pacific to determine the tail length when turtles mature.

```{r}
rm(list=ls())
library(tidyverse)
library(readr)
library(lubridate)
save.fig <- T

col_defs <- cols(Turtle_Number = col_integer(),
                 Trip = col_character(),
                 Set = col_character(),
                 DATE = col_integer(),
                 Latitude_DD = col_double(),
                 Longitude_DD = col_double(),
                 Sex = col_factor(levels = c("M", "F")),
                 Sex_1isF = col_integer(),
                 Maturity_State = col_factor(levels = c("Immature", "Mature")),
                 Species = col_skip(),
                 CCL_cm = col_double(),
                 SCL_cm = col_double(),
                 Total_Tail_cm = col_double(),
                 Plastron_to_Cloaca_cm = col_double(),
                 CaseAcc = col_character(),
                 Event_ID =col_character())

#, "Unknown")
#dat.CNP <- read_csv(file = "data/CNP_Lo_Bycatch_v2_May2022.csv",
dat.CNP <- read_csv(file = "data/CNP_Lo_Bycatch_v2_Dec2022.csv",
                    col_types = col_defs) %>%
  mutate(Date1 = ymd(DATE),
         Longitude = Longitude_DD,
         Latitude = Latitude_DD)
  
summary(dat.CNP)
```

```{r}
ggplot(dat.CNP) +
  geom_point(aes(x = Longitude, y = Latitude, color = Sex, size = CCL_cm))
```

```{r}
ggplot(dat.CNP) + 
  geom_point(aes(x = CCL_cm, y = SCL_cm, color = Maturity_State)) +
  facet_grid(rows = vars(Sex))
```

```{r}

juve <- dat.CNP %>% filter(Maturity_State == "Immature")
adults <- dat.CNP %>% filter(Maturity_State == "Mature")

c(max(juve$SCL_cm), min(adults$SCL_cm))
```

Look at how sex ratio changed over the years:

```{r}
dat.CNP %>% mutate(Year = year(Date1)) %>%
  group_by(Year) %>% summarise(N = n(), 
                               M = sum(Sex == "M"),
                               F = sum(Sex == "F")) %>%
  mutate(Groups = ifelse(Year < 2011, 1, 
                         ifelse(Year < 2016, 2, 3)))-> sex.by.year


sex.by.year %>% group_by(Groups) %>%
  summarize(N = sum(N), M = sum(M), F = sum(F)) %>%
  mutate(F_Ratio = F/N)-> tmp

tmp
```

Sex ratio didn't change that much over time (2005-2010 = `r signif(tmp[1,"F_Ratio"], 2)` (n=`r tmp[1,"N"]`), 2011-2015 = `r signif(tmp[2,"F_Ratio"], 2)` (n=`r tmp[2,"N"]`), 2016-`r max(sex.by.year$Year)` = `r signif(tmp[3,"F_Ratio"], 2)` (n=`r tmp[3,"N"]`))

Comparing SCL between locations...

Need to bring in the Epac dataset.


```{r}
col_def <- cols(ID = col_integer(),
                LabID = col_integer(),
                Date_Collected = col_date(),
                SCL_SWFSC = col_double(),
                CCL_SWFSC = col_double(),
                Predicted.Sex = col_character(),
                Mean.TotalTestosterone_pg_per_mL = col_double(),
                Duplicate1_TotalTestosterone_pg_per_mL = col_double(),
                Duplicate2_TotalTestosterone_pg_per_mL = col_double(),
                CaptureID = col_integer(),
                TH20_C = col_double(),
                Month = col_integer(),
                Day = col_integer(),
                Year = col_integer(),
                Time = col_skip(),   #col_time(format = "%H:%M"),
                Haplotype = col_character(),
                x = col_double(),
                y = col_double(),
                EEZ = col_character(),
                sex = col_character(),
                plastron_to_tip = col_double(),
                plastron_to_cloaca = col_double(),
                cloaca_to_tip = col_double(),
                #plastron_to_cloaca_to_tip = col_double(),
                Sum_plastron_to_cloaca_to_tip = col_double(),
                #dif_in_TL_measures = col_double(),
                Difference_columnU_columnX = col_double(),
                weight = col_double(),
                body_depth = col_double(),
                TP = col_double(),
                dN = col_double(),
                dC = col_double())

#file.name <- "data/EPacLO_2021-07-02.csv"
#file.name <- "data/EPac_LO_v2_Jan2022.csv"
#file.name <- "data/EPac_LO_v3_Jan2022.csv"
file.name <- "data/estimated_sex_LO_ETP_2022-06-14.csv"
dat.2 <- read_csv(file = file.name,
                  col_types = col_def)

summary(dat.2)

dat.2 %>% transmute(Date = Date_Collected,
                    SCL_cm = SCL_SWFSC,
                    CCL_cm = CCL_SWFSC,
                    Latitude = y,
                    Longitude = x,
                    Sex = as.factor(sex),
                    pMale = pMale,
                    Location = "ETP") %>%
  na.omit() -> dat.ETP

dat.CNP %>% transmute(Date = Date1,
                    SCL_cm = SCL_cm,
                    CCL_cm = CCL_cm,
                    Latitude = Latitude,
                    Longitude = Longitude,
                    Sex = Sex, 
                    pMale = ifelse(Sex == "M", 1, 0),
                    Location = "CNP") -> dat.CNP

dat.all <- rbind(dat.ETP, dat.CNP) %>%
  mutate(Est_Sex = ifelse(pMale > 0.5, "M", "F"))

ggplot(dat.all) +
  geom_histogram(aes(x = SCL_cm), binwidth = 5) + 
  facet_grid(Location ~ Est_Sex)

```

```{r}
ggplot(dat.all) +
  geom_point(aes(x = SCL_cm, y = CCL_cm)) + 
  facet_grid(Location ~ Est_Sex)

if (save.fig)
  ggsave(filename = "figures/CCL_SCL_Sex_location.png",
         device = "png", dpi = 600)
```

Fit Gaussian mixture models to ETP and CNP datasets separately.

Note that in mclust, BIC is defined as 2×ln(L(theta|x))+k×ln(n) not as −2×ln(L(theta|x))+k×ln(n). The latter is more often used. So, using mclust, we need to maximize BIC not to minimize. 

see these sites for more explanations:
https://github.com/data-edu/tidyLPA/issues/86
https://stats.stackexchange.com/questions/237220/mclust-model-selection


```{r}
library(mclust)
# compare one or two mixture models
CNP.mclust.BIC <- mclustBIC(dat.CNP$SCL_cm)
CNP.mclust <- densityMclust(dat.CNP$SCL_cm, x = CNP.mclust.BIC)

CNP.dens <- data.frame(SCL = CNP.mclust$data,
                       density = CNP.mclust$density) %>%
  arrange(SCL)

# Create histogram values:
CNP.hist <-  hist(CNP.dens$SCL, 
                  breaks = seq(from = 25, to = 70, by = 5), 
                  plot = F)

CNP.hist.df <- data.frame(breaks = CNP.hist$mids,
                          density = CNP.hist$density)

ggplot() + 
  geom_col(data = CNP.hist.df, aes(x = breaks, y = density),
           width = 5) +
  geom_path(data = CNP.dens, aes(x = SCL, y = density),
            size = 1) +
  labs(x = "SCL (cm)", y = "Density", title = "CNP") +
  theme(text = element_text(size = 12),
        axis.text = element_text(size = 11))
  #geom_path(aes(x = SCL, y = density)) +  
  #geom_rug(aes(x = SCL)) 

if (save.fig)
  ggsave(filename = "figures/CNP_SCL.png", device = "png", dpi = 600)
```

Two means for CNP are `r signif(CNP.mclust$parameters$mean, 4)` and their variances (`r signif(CNP.mclust$parameters$variance$sigmasq, 4)`).



```{r}
ETP.mclust.BIC <- mclustBIC(dat.ETP$SCL_cm)
ETP.mclust <- densityMclust(dat.ETP$SCL_cm, x = ETP.mclust.BIC)

ETP.dens <- data.frame(SCL = ETP.mclust$data,
                       density = ETP.mclust$density) %>%
  arrange(SCL)

# Create histogram values:
ETP.hist <-  hist(ETP.dens$SCL, 
                  breaks = seq(from = 15, to = 70, by = 5), 
                  plot = F)


ETP.hist.df <- data.frame(breaks = ETP.hist$mids,
                          density = ETP.hist$density)

ggplot() + 
  geom_col(data = ETP.hist.df, aes(x = breaks, y = density),
           width = 5) +
  geom_path(data = ETP.dens, aes(x = SCL, y = density),
            size = 1) +
    labs(x = "SCL (cm)", y = "Density", title = "ETP") +
  theme(text = element_text(size = 12),
        axis.text = element_text(size = 11))

if (save.fig)
  ggsave(filename = "figures/ETP_SCL.png", device = "png", dpi = 600)
```

So, for the ETP dataset, a mixture of 4 distributions maximizes BIC, whereas for the CNP dataset 2 distributions maximizes it. 


Also using sBIC package: (This is probably unnecessary)

```{r}
library(sBIC)

gMix <- GaussianMixtures(maxNumComponents = 4, phi = 1, restarts = 100) 

CNP.sBIC <- sBIC(dat.CNP$SCL_cm, gMix)
ETP.sBIC <- sBIC(dat.ETP$SCL_cm, gMix)
```

Try a simple KS test:

```{r}
KS.out <- ks.test(dat.CNP$SCL_cm, dat.ETP$SCL_cm, alternative = "two.sided")

KS.out
```

So... the distributions are different. Let's take a look at the eCDFs.

```{r}
SCLs <- data.frame(SCL = c(dat.CNP$SCL_cm, dat.ETP$SCL_cm),
                   Location = c(dat.CNP$Location, dat.ETP$Location))

ggplot(data = SCLs) +
  stat_ecdf(aes(x = SCL, color = Location))


```

The main difference is in the small size group < 35 cm. 

If we look at only those > 35 cm...

```{r}
SCLs %>% filter(SCL > 35) -> SCLs.big

KS.out.big <- ks.test(dat.CNP %>% filter(SCL_cm>35) %>% pull(SCL_cm), 
                      dat.ETP %>% filter(SCL_cm>35) %>% pull(SCL_cm), 
                      alternative = "two.sided")

KS.out.big
```

Even the big ones are different between the two locations. 


```{r}

ggplot(data = SCLs.big) +
  stat_ecdf(aes(x = SCL, color = Location))


```


```{r}
CNP.big.SCL <- dat.CNP %>% filter(SCL_cm > 35) %>% pull(SCL_cm)
CNP.big.mclust.BIC <- mclustBIC(CNP.big.SCL)
CNP.big.mclust <- densityMclust(CNP.big.SCL, x = CNP.big.mclust.BIC)

CNP.big.dens <- data.frame(SCL = CNP.big.mclust$data,
                       density = CNP.big.mclust$density) %>%
  arrange(SCL)

# Create histogram values:
CNP.big.hist <-  hist(CNP.big.SCL, 
                  breaks = seq(from = 35, to = 70, by = 5), 
                  plot = F)

CNP.big.hist.df <- data.frame(breaks = CNP.big.hist$mids,
                          density = CNP.big.hist$density)

ggplot() + 
  geom_col(data = CNP.big.hist.df, aes(x = breaks, y = density),
           width = 5) +
  geom_path(data = CNP.big.dens, aes(x = SCL, y = density),
            size = 1) +
  labs(x = "SCL (cm)", y = "Density", title = "CNP") +
  theme(text = element_text(size = 12),
        axis.text = element_text(size = 11))
  #geom_path(aes(x = SCL, y = density)) +  
  #geom_rug(aes(x = SCL)) 
```


```{r}
ETP.big.SCL <- dat.ETP %>% filter(SCL_cm > 35) %>% pull(SCL_cm)

ETP.big.mclust.BIC <- mclustBIC(ETP.big.SCL)
ETP.big.mclust <- densityMclust(ETP.big.SCL, x = ETP.big.mclust.BIC)

ETP.big.dens <- data.frame(SCL = ETP.big.mclust$data,
                       density = ETP.big.mclust$density) %>%
  arrange(SCL)

# Create histogram values:
ETP.big.hist <-  hist(ETP.big.dens$SCL, 
                  breaks = seq(from = 35, to = 70, by = 5), 
                  plot = F)


ETP.big.hist.df <- data.frame(breaks = ETP.big.hist$mids,
                          density = ETP.big.hist$density)

ggplot() + 
  geom_col(data = ETP.big.hist.df, aes(x = breaks, y = density),
           width = 5) +
  geom_path(data = ETP.big.dens, aes(x = SCL, y = density),
            size = 1) +
    labs(x = "SCL (cm)", y = "Density", title = "ETP") +
  theme(text = element_text(size = 12),
        axis.text = element_text(size = 11))
```




The following is not needed... 

Test the hypothesis that there is no difference between F/M and CNP/ETP, I use a randomization test to see how likely it is to see these minimum values if data were randomly placed in these groups. 


<!-- ```{r warning=FALSE} -->
<!-- n.rep <- 2500 -->
<!-- min.rnd <- matrix(nrow = n.rep, ncol = 4) -->
<!-- for (k in 1:n.rep){ -->
<!--   dat.all$SCL_rnd <- sample(dat.all$SCL_cm)  -->
<!--   dat.all %>% select(Location, Est_Sex, SCL_rnd) %>%  -->
<!--     group_by(Location, Est_Sex) %>% -->
<!--     summarize(min_SCL = min(SCL_rnd, na.rm = T)) -> tmp -->
<!--   min.rnd[k,] <- t(as.vector(tmp[, "min_SCL"])) -->

<!-- } -->


<!-- df.random <- data.frame(Location = rep(c("CNP", "ETP"), each = n.rep * 2), -->
<!--                         Sex = rep(rep(c("F", "M"), each = n.rep), 2), -->
<!--                         min_SCL = as.vector(min.rnd)) -->

<!-- # observed -->
<!-- dat.all %>% select(Location, Est_Sex, SCL_rnd)  %>%  -->
<!--   transmute(Location = Location, -->
<!--             Sex = Est_Sex, -->
<!--             SCL_rnd = SCL_rnd) %>% -->
<!--   group_by(Location, Sex) %>% -->
<!--     summarize(min_SCL = min(SCL_rnd, na.rm = T)) -> obs.SCL -->

<!-- ggplot() +  -->
<!--   geom_histogram(data = df.random, aes(x = min_SCL), binwidth = 2.5) + -->
<!--   geom_vline(data = obs.SCL, aes(xintercept = min_SCL), color = "red") +  -->
<!--   facet_grid(Location ~ Sex) -->

<!-- ``` -->



