---
title: "Analysis of olive ridleys in ETP"
output: html_notebook
---

Analysis of hormone data for olive ridleys in ETP. Feb 2022 version.

Testing how to use GitHub as a contributor rather than the primary. 

```{r}
library(jagsUI)
library(tidyverse)
library(lubridate)
library(readr)
library(loo)

save.fig <- F

compute.LOOIC <- function(loglik, data.vector, MCMC.params){
  n.per.chain <- (MCMC.params$n.samples - MCMC.params$n.burnin)/MCMC.params$n.thin
  
  loglik.vec <- as.vector(loglik)
  
  # each column corresponds to a data point and rows are MCMC samples
  loglik.mat <- matrix(loglik.vec, nrow = n.per.chain * MCMC.params$n.chains)
  
  # take out the columns that correspond to missing data points
  loglik.mat <- loglik.mat[, !is.na(data.vector)]
  # loglik.mat <- matrix(loglik.vec[!is.na(data.vector)], 
  #                      nrow = MCMC.params$n.chains * n.per.chain)
  
  Reff <- relative_eff(exp(loglik.mat),
                       chain_id = rep(1:MCMC.params$n.chains,
                                      each = n.per.chain),
                       cores = 4)
  
  #
  loo.out <- loo(loglik.mat, 
                 r_eff = Reff, 
                 cores = 4, k_threshold = 0.7)
  
  out.list <- list(Reff = Reff,
                   loo.out = loo.out)
  
  return(out.list)  
}

MCMC.params <- list(n.samples = 70000,
                    n.burnin = 50000,
                    n.thin = 5,
                    n.chains = 5)

```


```{r}
col_def <- cols(ID = col_integer(),
                LabID = col_integer(),
                Date_Collected = col_date(format = "%m/%d/%y"),
                SCL_SWFSC = col_double(),
                CCL_SWFSC = col_double(),
                Predicted.Sex = col_character(),
                Mean.TotalTestosterone_pg_per_mL = col_double(),
                Duplicate1_TotalTestosterone_pg_per_mL = col_double(),
                Duplicate2_TotalTestosterone_pg_per_mL = col_double(),
                CaptureID = col_integer(),
                TH20_C = col_double(),
                Month = col_integer(),
                Day = col_integer(),
                Year = col_integer(),
                Time = col_time(format = "%H:%M"),
                Haplotype = col_character(),
                x = col_double(),
                y = col_double(),
                EEZ = col_character(),
                sex = col_character(),
                plastron_to_tip = col_double(),
                plastron_to_cloaca = col_double(),
                cloaca_to_tip = col_double(),
                #plastron_to_cloaca_to_tip = col_double(),
                Sum_plastron_to_cloaca_to_tip = col_double(),
                #dif_in_TL_measures = col_double(),
                Difference_columnU_columnX = col_double(),
                weight = col_double(),
                body_depth = col_double(),
                TP = col_double(),
                dN = col_double(),
                dC = col_double())

#file.name <- "data/EPacLO_2021-07-02.csv"
#file.name <- "data/EPac_LO_v2_Jan2022.csv"
#file.name <- "data/EPac_LO_v3_Jan2022.csv"
file.name <- "data/EPac_LO_TH20_18Feb2022.csv"
dat.0 <- read_csv(file = file.name,
                  col_types = col_def)

summary(dat.0)
```

Filter necessary info:

```{r}
dat.0 %>% select(Mean.TotalTestosterone_pg_per_mL, 
                 Duplicate1_TotalTestosterone_pg_per_mL,
                 Duplicate2_TotalTestosterone_pg_per_mL, 
                 Date_Collected, sex,
                 plastron_to_tip, plastron_to_cloaca, cloaca_to_tip,
                 SCL_SWFSC, LabID, TH20_C,
                 weight, body_depth, dN, dC,
                 x, y, Haplotype,EEZ) %>%
  transmute(ID = LabID, 
            Mean.testo = Mean.TotalTestosterone_pg_per_mL,
            Testo1 = Duplicate1_TotalTestosterone_pg_per_mL,
            Testo2 = Duplicate2_TotalTestosterone_pg_per_mL,
            Date = Date_Collected,
            sex = sex,
            tail_length = plastron_to_tip,
            tail_1 = plastron_to_cloaca,
            tail_2 = cloaca_to_tip,
            SCL = SCL_SWFSC,
            log_Testo1_mean0 = log(Testo1),
            log_Testo2_mean0 = log(Testo2),
            sex01 = ifelse(sex == "F", 0,
                           ifelse(sex == "M", 1, NA)),
            weight = weight,
            body_depth = body_depth,
            dN = dN,
            dC = dC,
            lat = y,
            lon = x,
            Haplotype = as.factor(Haplotype),
            EEZ = as.factor(EEZ), 
            TempC = TH20_C) -> dat.1

summary(dat.1)
```

Take a look at some relationships among variables. 


```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = log(Mean.testo), y = tail_length, color = sex)) 
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = log(Mean.testo), y = SCL, color = sex))
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = log(Mean.testo), y = body_depth, color = sex)) 
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = log(Mean.testo)))  
```

```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = sex))  

```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = dN))
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = dC))

```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = dN/dC))

```



```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = SCL))

```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = SCL, y = weight, color = sex))  

if (save.fig)
  ggsave(filename = "figures/SCL_weight.png",
         device = "png", dpi = 400)
```

For missing values in temperature, latitude ranged from 14.92 to 16.57 and longitude from -100.52 to -99.38. 

I selected all data in this lat/lon range and gave the mean value to the missing values.

```{r}

dat.1 %>% filter(is.na(TempC)) %>%
  select(lat, lon) %>% 
  summarise(min.lat = min(lat),
            max.lat = max(lat),
            min.lon = min(lon),
            max.lon = max(lon)) -> lat.lon.TempEqNA

dat.1 %>% filter(lat >= lat.lon.TempEqNA$min.lat & 
                   lat <= lat.lon.TempEqNA$max.lat &
                   lon >= lat.lon.TempEqNA$min.lon &
                   lon <= lat.lon.TempEqNA$max.lat) %>%
  summarise(mean.temp = mean(TempC, na.rm = T),
            sd.temp = sqrt(var(TempC, na.rm = T))) -> mean.Temp

dat.1$TempC[is.na(dat.1$TempC)] <- as.double(mean.Temp$mean.temp)

ggplot(data = dat.1 %>% filter()) +
  geom_point(aes(x = lat, y = TempC))
```


We decided to use the regression line to estimate SCL of 65833, which has weight but not SCL. 

```{r}
ggplot(data = dat.1) +
  geom_point(aes(y = SCL, x = log(weight), color = sex))


```


```{r}
lm_WvsSCL <- lm(SCL ~ log(weight) , data = dat.1)
summary(lm_WvsSCL)

```

Predict SCL of 65833

```{r}
dat.1 %>% filter(ID == 65833) %>% select(weight) %>% as.numeric -> weight_65833

fit_65833 <- predict(lm_WvsSCL, newdata = data.frame(weight = weight_65833),
        interval = "confidence")
#log(weight_65833) = 3.35
dat.1 %>% 
  mutate(SCL = replace(SCL, ID == 65833, fit_65833[1])) -> dat.1

```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = SCL, y = tail_length/SCL, color = sex))  

if (save.fig)
  ggsave(filename = "figures/SCL_weight_Feb2022.png",
         device = "png", dpi = 400)
```


```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = tail_length, y = tail_1)) +
  geom_point(data = dat.1,
             aes(x = tail_length, y = tail_2),
             color = "orange") +
  geom_point(data = dat.1,
             aes(x = tail_length, y = tail_1 + tail_2),
             color = "red") +
  labs(y = "tail measurements (cm)", x = "total tail length (cm)")

```

There are different variability in three tail length measurements. Find the best one (or the least variable).


```{r}
lm1 <- lm(tail_length ~ tail_1, data = dat.1)
lm2 <- lm(tail_length ~ tail_2, data = dat.1)
dat.1 %>% mutate(tail_12 = tail_1 + tail_2) -> dat.1.1
lm3 <- lm(tail_length ~ tail_12, data = dat.1.1)

```

Interestingly, tail_1 + tail_2 (plastron to cloaca and cloaca to tip) don't add up to total length. Tail_2 (cloaca to tip) seems to be the least variable measurement. 


```{r}
models <- c("sex_only", 
            "sex_tail", "sex_tail", "sex_temp",
            "sex_tail_SCL", "sex_tail_SCL", "sex_tail_temp", "sex_tail_temp",
            "sex_tail_SCL_temp", "sex_tail_SCL_temp",
            "sex_tail", "sex_tail", 
            "sex_tail_SCL", "sex_tail_SCL", "sex_tail_temp", "sex_tail_temp",
            "sex_tail_SCL_temp", "sex_tail_SCL_temp")

model.names <- c("sex_only", 
                 "sex_tail", "sex_tail2", "sex_temp",
                 "sex_tail_SCL", "sex_tail2_SCL", "sex_tail_temp", "sex_tail2_temp",
                 "sex_tail_SCL_temp", "sex_tail2_SCL_temp",
                 "sex_ptail", "sex_ptail2", 
                 "sex_ptail_SCL", "sex_ptail2_SCL", "sex_ptail_temp", "sex_ptail2_temp",
                 "sex_ptail_SCL_temp", "sex_ptail2_SCL_temp")

# use mean for missing tail length:
dat.1$tail_1[is.na(dat.1$tail_1)] <- mean(dat.1$tail_1, na.rm = T)

# use mean for missing tail length:
dat.1$tail_2[is.na(dat.1$tail_2)] <- mean(dat.1$tail_2, na.rm = T)

# use the sum of tail_1 and tail_2 for missing tail length:
dat.1$tail_length[is.na(dat.1$tail_length)] <- dat.1$tail_1[is.na(dat.1$tail_length)] + dat.1$tail_2[is.na(dat.1$tail_length)] 

# also use mean for missing SCL
dat.1$SCL[is.na(dat.1$SCL)] <- mean(dat.1$SCL, na.rm = T)


jags.data.list <- list(list(H1 = dat.1[, c("log_Testo1_mean0",    #sex_only
                                           "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1)),
                    list(H1 = dat.1[, c("log_Testo1_mean0",       #sex_tail
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length),
                    list(H1 = dat.1[, c("log_Testo1_mean0",      # sex_tail2
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2),
                    list(H1 = dat.1[, c("log_Testo1_mean0",      # sex_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         temp = dat.1$TempC),
                    list(H1 = dat.1[, c("log_Testo1_mean0",    # sex_tail_SCL
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_tail2_SCL
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_tail_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length,
                         temp = dat.1$TempC),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_tail2_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2,
                         temp = dat.1$TempC),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_tail_SCL_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length,
                         temp = dat.1$TempC,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_tail2_SCL_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2,
                         temp = dat.1$TempC,
                         SCL = dat.1$SCL),
                    
                    list(H1 = dat.1[, c("log_Testo1_mean0",    #sex_ptail
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length/dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_ptail2
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2/dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   # sex_ptail_SCL
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length/dat.1$SCL,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   # sex_ptail2_SCL
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2/dat.1$SCL,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_ptail_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length/dat.1$SCL,
                         temp = dat.1$TempC),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_ptail2_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2/dat.1$SCL,
                         temp = dat.1$TempC),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_ptail_SCL_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length/dat.1$SCL,
                         temp = dat.1$TempC,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0",   #sex_ptail2_SCL_temp
                                        "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2/dat.1$SCL,
                         temp = dat.1$TempC,
                         SCL = dat.1$SCL))

parameters <- c("Sex1", 
                "beta_0", "beta_Sex", "beta_T", "beta_L", "beta_temp",
                "mu_H1", "sigma_H1", 
                "deviance", "loglik")


```


```{r}


k <- 1
LOOIC <- DIC <- vector(mode = "numeric", length = length(models))

for (k in 1:length(models)){
  
  if (!file.exists(paste0("RData/LO_", model.names[[k]],
                          "_Feb2022.rds"))){
    jags.data <- jags.data.list[[k]]
    MCMC.params$parameters <- parameters
    MCMC.params$model.file <- paste0("models/Model_", models[[k]], ".txt")
    jm.sex <- jags(data = jags.data,
                   #inits = inits,
                   parameters.to.save= MCMC.params$parameters,
                   model.file = MCMC.params$model.file,
                   n.chains = MCMC.params$n.chains,
                   n.burnin = MCMC.params$n.burnin,
                   n.thin = MCMC.params$n.thin,
                   n.iter = MCMC.params$n.samples,
                   DIC = T, 
                   parallel=T)
    
    jm.out <- list(data = jags.data,
                   params = MCMC.params,
                   jm.out = jm.sex)
    
    saveRDS(jm.out, file = paste0("RData/LO_", model.names[[k]],
                                  "_Feb2022.rds"))
  } else {
    jm.out <- readRDS(file = paste0("RData/LO_", model.names[[k]],
                                    "_Feb2022.rds"))
    
  }
  
  DIC[k] <- jm.out$jm.out$DIC
  if (!file.exists(paste0("RData/LO_", model.names[[k]],
                          "_loo_out_Feb2022.rds"))){
    loo.out <- compute.LOOIC(loglik = jm.out$jm.out$sims.list$loglik,
                             data.vector = as.vector(jm.out$data$H1),
                             MCMC.params = jm.out$params)
    
    saveRDS(loo.out, paste0("RData/LO_", model.names[[k]], 
                            "_loo_out_Feb2022.rds") ) 
  } else {
    loo.out <- readRDS(paste0("RData/LO_", model.names[[k]],
                              "_loo_out_Feb2022.rds"))
  }
  LOOIC[k] <- loo.out$loo.out$estimates["looic","Estimate"]
}


```




Compare results:
```{r}

LOOIC_DIC <- data.frame(Model = model.names,
             dDIC = DIC - min(DIC),
             dLOOIC = LOOIC - min(LOOIC)) %>% arrange(by = dLOOIC)
LOOIC_DIC
```

Write out the LOOIC table
```{r}

write_csv(LOOIC_DIC, file = "data/model_comp_Feb2022.csv")

```


```{r}
best.model.name <- LOOIC_DIC[1, "Model"] # "sex_ptail_SCL"
best.model <- readRDS(paste0("RData/LO_", 
                             best.model.name, "_Feb2022.rds"))
dat.1$Model <- best.model.name
dat.1$pMale <- best.model$jm.out$mean$Sex1
dat.1$log_muH <- best.model$jm.out$q50$mu_H1

dat.1 %>% select(ID, tail_1, tail_2, 
                 log_Testo1_mean0, 
                 log_Testo2_mean0, pMale) %>%
  mutate(LabID = ID) %>%  
  select(-ID) %>% 
  right_join(dat.0, by = "LabID") -> dat.2.ptail

loo.out <- readRDS(paste0("RData/LO_", best.model.name, 
                          "_loo_out_Feb2022.rds") )
plot(loo.out$loo.out)
```

There are some high Pareto K statistics... Take a look at the second best model:

```{r}
second.model.name <- LOOIC_DIC[2, "Model"] 
second.model <- readRDS(paste0("RData/LO_", 
                             second.model.name, "_Feb2022.rds"))
dat.1$Model <- second.model.name
dat.1$pMale <- second.model$jm.out$mean$Sex1
dat.1$log_muH <- second.model$jm.out$q50$mu_H1

loo.out <- readRDS(paste0("RData/LO_", second.model.name, 
                          "_loo_out_Feb2022.rds") )
plot(loo.out$loo.out)


```

About the same as the first one. 

```{r}
third.model.name <- LOOIC_DIC[3, "Model"] 
third.model <- readRDS(paste0("RData/LO_", 
                             third.model.name, "_Feb2022.rds"))
dat.1$Model <- third.model.name
dat.1$pMale <- third.model$jm.out$mean$Sex1
dat.1$log_muH <- third.model$jm.out$q50$mu_H1

loo.out <- readRDS(paste0("RData/LO_", third.model.name, 
                          "_loo_out_Feb2022.rds") )
plot(loo.out$loo.out)

```

Although the sex_ptail_SCL and sex_ptail_SCL_temp models are the first and secon best, PSIS diagnostic plot showed more high values (>0.7) of Pareto k statistics than the sex_tail_SCL model. So, I decided to use the sex_tail_SCL.  


Only 3 data points showed high Pareto k values (> 0.7) whereas others were all < 0.7. So, I'll just go with this model. Look at the results:

```{r}
# combine the results with dat.0
dat.1 %>% select(ID, tail_1, tail_2, 
                 log_Testo1_mean0, 
                 log_Testo2_mean0, pMale) %>%
  mutate(LabID = ID) %>%  
  select(-ID) %>% 
  right_join(dat.0, by = "LabID") -> dat.2
#
if (!file.exists(paste0("data/estimated_sex_LO_ETP_", 
                        Sys.Date(), ".csv")))
  write.csv(dat.2,
            file = paste0("data/estimated_sex_LO_ETP_", 
                          Sys.Date(), ".csv"),
            quote = FALSE, row.names = FALSE)

# Look at how pMale were similar between the two models.
summary(lm(dat.2$pMale ~ dat.2.ptail$pMale))

```

Take a look at the high Pareto k value data points:

```{r}
loo.out$loo.out$pointwise %>% 
  data.frame() %>%
  rownames_to_column() -> loo.out.df

loo.out.df %>% filter(influence_pareto_k > 0.7) -> high_pareto_k

high_pareto_k <- high_pareto_k %>% 
  mutate(rowID = as.numeric(rowname)) %>%
  mutate(rowID = replace(rowID, rowID > nrow(dat.1), rowID[rowID > nrow(dat.1)] - nrow(dat.1))) 

#dat.1 %>% mutate(var_Testo = ((Testo1 - Mean.testo)^2 + (Testo2 - Mean.testo)^2)) -> dat.1

```


```{r}
dat.1 %>% mutate(log_Testo_mean = (log_Testo1_mean0 + log_Testo2_mean0)/2,
                 log_Testo_var = (log_Testo1_mean0 - log_Testo_mean)^2 + 
                   (log_Testo2_mean0 - log_Testo_mean)^2) -> dat.1

ggplot() + geom_point(data = dat.1,
                      aes(x = ID, y = log_Testo_var)) +
  geom_point(data = dat.1[high_pareto_k$rowID,],
             aes(x = ID, y = log_Testo_var),
             color = "red")
```


```{r}
dat.1[high_pareto_k$rowID, c("ID", "Mean.testo", "log_Testo_mean", "SCL", "tail_length")]
```


```{r}

ggplot(data = dat.1) +
  geom_point(aes(x = SCL, y = tail_length, color = sex)) +
  geom_point(data = dat.1 %>% filter(ID == 65937 | ID == 65844 | ID == 65856),
             aes(x = SCL, y = tail_length),
             shape = 1, size = 2)


```

Not sure why these three came back with high Pareto K values...


Sex ratio of the ETP data
```{r}

dat.2 %>% mutate(Estimated_Sex = ifelse(pMale > 0.5, 
                                        "Male", "Female")) %>% 
  group_by(Estimated_Sex) %>% 
  summarize(n = n(), ratio = n/nrow(dat.2)) -> Sex.ratio

dat.2 %>% mutate(Estimated_Sex = ifelse(pMale > 0.5, 
                                        "Male", "Female")) %>% 
  filter(sex == "I") %>%
  group_by(Estimated_Sex) %>% 
  summarize(n = n(), ratio = n/nrow(dat.2)) -> Sex.ratio.immature

dat.2 %>% mutate(Estimated_Sex = ifelse(pMale > 0.5, 
                                        "Male", "Female")) %>% 
  filter(sex != "I") %>%
  group_by(Estimated_Sex) %>% 
  summarize(n = n(), ratio = n/nrow(dat.2)) -> Sex.ratio.mature
```




make some plots:

```{r}
ggplot() + 
  geom_point(data = dat.1,
             aes(x = ID, y = pMale,
                 size = log(Mean.testo),
                 color = as.factor(sex01)))
```


```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = log_muH, y = pMale,
                 color = as.factor(sex01)))

# ggsave(filename = "figures/log_muH_Vs_pMale.png",
#        device = "png", dpi = 600)
```



```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = tail_length, y = pMale,
                 color = as.factor(sex01)))
```




```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = SCL, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = tail_length/SCL, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
dat.1 %>% mutate(estimatedSex = ifelse(pMale>0.5, "Male", "Female")) -> dat.1

ggplot(data = dat.1) +
  geom_point(aes(x = SCL, y = tail_length/SCL, color = estimatedSex))

```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = SCL, y = tail_length, color = estimatedSex))

```



```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = body_depth, y = pMale,
                 color = as.factor(sex01)))

```


```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = dN, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = dC, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = dN/dC, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
dat.haplo <- dat.1 %>% 
  group_by(Haplotype) %>% 
  summarise(n = length(Haplotype), 
            mean.pMale = mean(pMale))

ggplot() +
  geom_boxplot(data = dat.1,
             aes(x = Haplotype, y = pMale)) +
  geom_text(data = dat.haplo,
            aes(x = Haplotype, y = 1.2, label = n)) +
  geom_text(data = dat.haplo,
            aes(x = Haplotype, y = 1.1, label = signif(mean.pMale, 1)))

```

