---
title: "Analysis of olive ridleys in ETP"
output: html_notebook
---

Analysis of hormone data for olive ridleys in ETP

```{r}
library(jagsUI)
library(tidyverse)
library(lubridate)
library(readr)
library(loo)

compute.LOOIC <- function(loglik, data.vector, MCMC.params){
  n.per.chain <- (MCMC.params$n.samples - MCMC.params$n.burnin)/MCMC.params$n.thin
  
  loglik.vec <- as.vector(loglik)
  
  # each column corresponds to a data point and rows are MCMC samples
  loglik.mat <- matrix(loglik.vec, nrow = n.per.chain * MCMC.params$n.chains)
  
  # take out the columns that correspond to missing data points
  loglik.mat <- loglik.mat[, !is.na(data.vector)]
  # loglik.mat <- matrix(loglik.vec[!is.na(data.vector)], 
  #                      nrow = MCMC.params$n.chains * n.per.chain)
  
  Reff <- relative_eff(exp(loglik.mat),
                       chain_id = rep(1:MCMC.params$n.chains,
                                      each = n.per.chain),
                       cores = 4)
  
  #
  loo.out <- loo(loglik.mat, 
                 r_eff = Reff, 
                 cores = 4, k_threshold = 0.7)
  
  out.list <- list(Reff = Reff,
                   loo.out = loo.out)
  
  return(out.list)  
}

MCMC.params <- list(n.samples = 50000,
                    n.burnin = 30000,
                    n.thin = 5,
                    n.chains = 5)

```


```{r}
col_def <- cols(ID = col_integer(),
                LabID = col_integer(),
                Date_Collected = col_date(format = "%m/%d/%Y"),
                SCL_SWFSC = col_double(),
                CCL_SWFSC = col_double(),
                Predicted.Sex = col_character(),
                Mean.TotalTestosterone_pg_per_mL = col_double(),
                Duplicate1_TotalTestosterone_pg_per_mL = col_double(),
                Duplicate2_TotalTestosterone_pg_per_mL = col_double(),
                CaptureID = col_integer(),
                Month = col_integer(),
                Day = col_integer(),
                Year = col_integer(),
                Time = col_time(format = "%H:%M"),
                Haplotype = col_character(),
                x = col_double(),
                y = col_double(),
                EEZ = col_character(),
                sex = col_character(),
                plastron_to_tip = col_double(),
                plastron_to_cloaca = col_double(),
                cloaca_to_tip = col_double(),
                plastron_to_cloaca_to_tip = col_double(),
                dif_in_TL_measures = col_double(),
                weight = col_double(),
                body_depth = col_double(),
                TP = col_double(),
                dN = col_double(),
                dC = col_double())

dat.0 <- read_csv(file = "data/EPacLO_2021-06-25.csv",
                  col_types = col_def)

summary(dat.0)
```

Filter necessary info:

```{r}
dat.0 %>% select(Mean.TotalTestosterone_pg_per_mL, 
                 Duplicate1_TotalTestosterone_pg_per_mL,
                 Duplicate2_TotalTestosterone_pg_per_mL, 
                 Date_Collected, sex,
                 plastron_to_tip, plastron_to_cloaca, cloaca_to_tip,
                 SCL_SWFSC, LabID,
                 weight, body_depth, dN, dC,
                 x, y, Haplotype,EEZ) %>%
  transmute(ID = LabID, 
            Mean.testo = Mean.TotalTestosterone_pg_per_mL,
            Testo1 = Duplicate1_TotalTestosterone_pg_per_mL,
            Testo2 = Duplicate2_TotalTestosterone_pg_per_mL,
            Date = Date_Collected,
            sex = sex,
            tail_length = plastron_to_tip,
            tail_1 = plastron_to_cloaca,
            tail_2 = cloaca_to_tip,
            SCL = SCL_SWFSC,
            log_Testo1_mean0 = log(Testo1),
            log_Testo2_mean0 = log(Testo2),
            sex01 = ifelse(sex == "F", 0,
                           ifelse(sex == "M", 1, NA)),
            weight = weight,
            body_depth = body_depth,
            dN = dN,
            dC = dC,
            lat = y,
            lon = x,
            Haplotype = as.factor(Haplotype),
            EEZ = as.factor(EEZ)) -> dat.1

summary(dat.1)
```

Take a look at some relationships among variables

```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = log(Mean.testo), y = tail_length, color = sex)) +
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = log(Mean.testo), y = tail_length),
             shape = 1, size = 2)
```

The one with black circle (ID == 65855) has high probability (~0.25) of male with low testosterone concentration (see plots below).

```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = log(Mean.testo), y = SCL, color = sex))+
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = log(Mean.testo), y = SCL),
             shape = 1, size = 2)
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = log(Mean.testo), y = body_depth, color = sex)) +
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = log(Mean.testo), y = body_depth),
             shape = 1, size = 2)
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = log(Mean.testo)))  +
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = lon, y = lat),
             shape = 1, size = 2, color = "orange")
```

```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = sex))  +
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = lon, y = lat),
             shape = 1, size = 2, color = "black")
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = dN))
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = dC))+
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = lon, y = lat),
             shape = 1, size = 2, color = "orange")
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = dN/dC))+
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = lon, y = lat),
             shape = 1, size = 2, color = "orange")
```



```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = lon, y = lat, color = SCL))+
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = lon, y = lat),
             shape = 1, size = 2, color = "orange")
```


```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = SCL, y = weight, color = sex))  +
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = SCL, y = weight),
             shape = 1, size = 2, color = "black")

# ggsave(filename = "figures/SCL_weight.png",
#        device = "png", dpi = 400)
```



```{r}
ggplot(data = dat.1) +
  geom_point(aes(x = SCL, y = tail_length/SCL, color = sex))  +
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = SCL, y = tail_length/SCL),
             shape = 1, size = 2, color = "black")

# ggsave(filename = "figures/SCL_weight.png",
#        device = "png", dpi = 400)
```


```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = tail_length, y = tail_1)) +
  geom_point(data = dat.1,
             aes(x = tail_length, y = tail_2),
             color = "orange") +
  geom_point(data = dat.1,
             aes(x = tail_length, y = tail_1 + tail_2),
             color = "red")  +
  geom_point(data = dat.1 %>% filter(ID == 65855),
             aes(x = tail_length, y = tail_1),
             shape = 1, size = 2, color = "green") +
  labs(y = "tail measurements (cm)", x = "total tail length (cm)")

```

There are different variability in three tail length measurements. Find the best one (or the least variable).


```{r}
lm1 <- lm(tail_length ~ tail_1, data = dat.1)
lm2 <- lm(tail_length ~ tail_2, data = dat.1)
dat.1 %>% mutate(tail_12 = tail_1 + tail_2) -> dat.1.1
lm3 <- lm(tail_length ~ tail_12, data = dat.1.1)

```

Interestingly, tail_1 + tail_2 (plastron to cloaca and cloaca to tip) don't add up to total length. Tail_2 (cloaca to tip) seems to be the least variable measurement. 


```{r}
models <- c("sex_only", "sex_tail", "sex_tail", 
            "sex_tail_SCL", "sex_tail_SCL",
            "sex_tail", "sex_tail", 
            "sex_tail_SCL", "sex_tail_SCL")

model.names <- c("sex_only", "sex_tail", "sex_tail2", 
                 "sex_tail_SCL", "sex_tail2_SCL",
                 "sex_ptail", "sex_ptail2", 
                 "sex_ptail_SCL", "sex_ptail2_SCL")

# use mean for missing tail length:
dat.1$tail_1[is.na(dat.1$tail_1)] <- mean(dat.1$tail_1, na.rm = T)

# use mean for missing tail length:
dat.1$tail_2[is.na(dat.1$tail_2)] <- mean(dat.1$tail_2, na.rm = T)

# use the sum of tail_1 and tail_2 for missing tail length:
dat.1$tail_length[is.na(dat.1$tail_length)] <- dat.1$tail_1[is.na(dat.1$tail_length)] + dat.1$tail_2[is.na(dat.1$tail_length)] 

# also use mean for missing SCL
dat.1$SCL[is.na(dat.1$SCL)] <- mean(dat.1$SCL, na.rm = T)

jags.data.list <- list(list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1)),
                    list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length),
                    list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2),
                    list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length/dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2/dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_length/dat.1$SCL,
                         SCL = dat.1$SCL),
                    list(H1 = dat.1[, c("log_Testo1_mean0", "log_Testo2_mean0")],
                         Sex1 = dat.1$sex01,
                         N1 = nrow(dat.1),
                         tail = dat.1$tail_2/dat.1$SCL,
                         SCL = dat.1$SCL))

parameters.list <- list(c("Sex1", "beta_0", "beta_Sex", 
                       "mu_H1", "sigma_H1", "deviance",
                       "loglik"),
                     c("Sex1", "beta_0", 
                       "beta_Sex", "beta_T",
                       "mu_H1", "sigma_H1", 
                       "deviance", "loglik"),
                     c("Sex1", "beta_0", 
                       "beta_Sex", "beta_T",
                       "mu_H1", "sigma_H1", 
                       "deviance", "loglik"),
                     c("Sex1", "beta_0", "beta_Sex", 
                       "beta_T", "beta_L",
                       "mu_H1", "sigma_H1", 
                       "deviance", "loglik"),
                     c("Sex1", "beta_0", "beta_Sex", 
                       "beta_T", "beta_L",
                       "mu_H1", "sigma_H1", 
                       "deviance", "loglik"),
                     c("Sex1", "beta_0", 
                       "beta_Sex", "beta_T",
                       "mu_H1", "sigma_H1", 
                       "deviance", "loglik"),
                     c("Sex1", "beta_0", 
                       "beta_Sex", "beta_T",
                       "mu_H1", "sigma_H1", 
                       "deviance", "loglik"),
                     c("Sex1", "beta_0", 
                       "beta_Sex", "beta_T",
                       "mu_H1", "sigma_H1", 
                       "deviance", "loglik"),
                     c("Sex1", "beta_0", 
                       "beta_Sex", "beta_T",
                       "mu_H1", "sigma_H1", 
                       "deviance", "loglik") )

```


```{r}


k <- 1
LOOIC <- DIC <- vector(mode = "numeric", length = length(models))

for (k in 1:length(models)){
  
  if (!file.exists(paste0("RData/LO_", model.names[[k]], ".rds"))){
    jags.data <- jags.data.list[[k]]
    MCMC.params$parameters <- parameters.list[[k]]
    MCMC.params$model.file <- paste0("models/Model_", models[[k]], ".txt")
    jm.sex <- jags(data = jags.data,
                   #inits = inits,
                   parameters.to.save= MCMC.params$parameters,
                   model.file = MCMC.params$model.file,
                   n.chains = MCMC.params$n.chains,
                   n.burnin = MCMC.params$n.burnin,
                   n.thin = MCMC.params$n.thin,
                   n.iter = MCMC.params$n.samples,
                   DIC = T, 
                   parallel=T)
    
    jm.out <- list(data = jags.data,
                   params = MCMC.params,
                   jm.out = jm.sex)
    
    saveRDS(jm.out, file = paste0("RData/LO_", model.names[[k]], ".rds"))
  } else {
    jm.out <- readRDS(file = paste0("RData/LO_", model.names[[k]], ".rds"))
    
  }
  
  DIC[k] <- jm.out$jm.out$DIC
  if (!file.exists(paste0("RData/LO_", model.names[[k]], "_loo_out.rds"))){
    loo.out <- compute.LOOIC(loglik = jm.out$jm.out$sims.list$loglik,
                             data.vector = as.vector(jm.out$data$H1),
                             MCMC.params = jm.out$params)
    
    saveRDS(loo.out, paste0("RData/LO_", model.names[[k]], "_loo_out.rds") ) 
  } else {
    loo.out <- readRDS(paste0("RData/LO_", model.names[[k]], "_loo_out.rds"))
  }
  LOOIC[k] <- loo.out$loo.out$estimates["looic","Estimate"]
}


```




Compare results:
```{r}

LOOIC_DIC <- data.frame(Model = model.names,
             dDIC = DIC - min(DIC),
             dLOOIC = LOOIC - min(LOOIC)) %>% arrange(by = dLOOIC)
LOOIC_DIC
```

The sex_ptail_SCL model is best.  We should also use Pareto k statistic to see if the model fit right. 

```{r}
best.model.name <- "sex_ptail_SCL"
best.model <- readRDS(paste0("RData/LO_", best.model.name, ".rds"))
dat.1$Model <- "sex_ptail_SCL"
dat.1$pMale <- best.model$jm.out$mean$Sex1
dat.1$log_muH <- best.model$jm.out$q50$mu_H1

loo.out <- readRDS(paste0("RData/LO_", best.model.name, "_loo_out.rds") )
plot(loo.out$loo.out)
```

Only two data points showed high Pareto k values (> 0.7) whereas others were all < 0.7. So, I'll just go with this model. Look at the results:

```{r}
# combine the results with dat.0
dat.1 %>% select(ID, tail_1, tail_2, log_Testo1_mean0, log_Testo2_mean0, pMale) %>%
  mutate(LabID = ID) %>%  select(-ID) %>% right_join(dat.0, by = "LabID") -> dat.2
#
#if (!file.exists("data/estimated_sex_LO_ETP.csv"))
  write.csv(dat.2, 
            file = paste0("data/estimated_sex_LO_ETP_", Sys.Date(), ".csv"),
            quote = FALSE, row.names = FALSE)
```

make some plots:

```{r}
ggplot() + 
  geom_point(data = dat.1,
             aes(x = ID, y = pMale,
                 size = log(Mean.testo),
                 color = as.factor(sex01)))
```

Find the ID of somewhat high p(male) with low Testo values. 

```{r}
dat.1 %>% filter(pMale > 0.25 & pMale < 0.5)
```



```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = log_muH, y = pMale,
                 color = as.factor(sex01)))

# ggsave(filename = "figures/log_muH_Vs_pMale.png",
#        device = "png", dpi = 600)
```



```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = tail_length, y = pMale,
                 color = as.factor(sex01)))
```




```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = SCL, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = tail_length/SCL, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = body_depth, y = pMale,
                 color = as.factor(sex01)))

```


```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = dN, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = dC, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
ggplot() +
  geom_point(data = dat.1,
             aes(x = dN/dC, y = pMale,
                 color = as.factor(sex01)))

```

```{r}
dat.haplo <- dat.1 %>% 
  group_by(Haplotype) %>% 
  summarise(n = length(Haplotype), 
            mean.pMale = mean(pMale))

ggplot() +
  geom_boxplot(data = dat.1,
             aes(x = Haplotype, y = pMale)) +
  geom_text(data = dat.haplo,
            aes(x = Haplotype, y = 1.2, label = n)) +
  geom_text(data = dat.haplo,
            aes(x = Haplotype, y = 1.1, label = signif(mean.pMale, 1)))

```

